Conflict Driven Evolution and Urban Development: \\ Evidence from the Holy Roman Empire\footnote{ I thank Lara Andermann, Leonie Bielefeld, Daniel Bursian, Davide Cantoni, Emilio Esguerra, Filip Milojevic, Johannes Vorbach and Matthias Weigand for helpful comments and feedback. }

### 2023-06-26 paper/paper.tex
#29 finish spatial heterogeneity part

\cite{diamond1997} and \cite{landes1969, landes2006} have argued for the economic benefits of political fragmentation and state competition. \cite{diamond1997} asserts that geography 
remoteness of China deprived it of the benefits of state competition that European states enjoyed throughout their history. \cite{landes2006} makes a similar argument, pointing out how the totalitarianism and bureaucratism that stifled the spread of innovation in medieval China would not have survived long in a more fragmented state system. Fragmentation also plays a central role in \cite{cervellati2022}'s theory.

The closest relative to this paper is \cite{schoenholzer2022} who investigate how switching between states affects the development of cities. They find that switching states affects population negatively in the short run but positively in the long run. Further, they show that this long-run benefit is driven by improved state quality after switching. Similar to the evolution metaphor, they interpret this as "creative destruction": the short run cost of destruction is compensated by the long-run benefit of becoming part of a higher-quality state. However, they do not show explicitly how these improvements in state quality are driven by selection through conflict. They use a dataset with a bigger geographical scope (all of Europe) but lower resolution: they use urban population from \cite{bairoch1988} as a proxy for development, which is only available in intervals of 50 or 100 years.

@@

While at first this result may seem like evidence for omitted variable bias in the baseline estimate --- the "correct" estimate being the negative one in Column (3) --- a closer look at the coefficients on $Conflict$ makes this seem unlikely: they are close to zero and not significant by any stretch of the imagination. To cause bias conflict incidents would need to be correlated both with the post-$Conquest$ dummy and with construction, but the correlation with construction is empirically zero.\footnote
{
The coefficients on $Conflict$ in Table \ref{tab:controls_50y} are conditional correlations. Simply computing the correlation coefficient between $Conflict$ and $All\,construction$ yields a value of $-0.01$.
} 
Rather, I suggest that the coefficients are different because they are estimated on different samples: more than two thirds of observations are lost between Columns (2) and (3) because of missing values in the $Conflict$ variable. Figure \ref{fig:conflict_map} shows that the remaining observations follow quite a different distribution across space: The colour of each region codes the share of observations from the baseline sample that are included in the conflict sample. If the spatial composition of the sample stayed exactly the same, the entire map would be a uniform shade. Instead we immediately see big discrepancies: Thuringia and most northern regions all but disappear from the sample, while the more darkly shaded west and south retain between 30\% and 60\% of their observations.

If the difference between the coefficients in (1) and (3) is due to the different spatial distribution of the sample, that has the interesting implication of spatially heterogeneous effects. BLANK.

>>>>

\cite{diamond1997} and \cite{landes1969, landes2006} have argued for the economic benefits of political fragmentation and state competition. \cite{diamond1997} asserts that geography 
remoteness of China deprived it of the benefits of state competition that European states enjoyed throughout their history. \cite{landes2006} makes a similar argument, pointing out how the totalitarianism and bureaucratism that stifled the spread of innovation in medieval China would not have survived long in a more fragmented state system. Geography and fragmentation also play central roles in \cite{cervellati2022}'s theory of European state history.

The closest relative to this paper is \cite{schoenholzer2022} who investigate how switching between states affects the development of cities. They find that switching states affects population negatively in the short run but positively in the long run. Further, they show that this long-run benefit is driven by improved state quality after switching. Similar to the evolution metaphor, they interpret this as "creative destruction": the short run cost of destruction is compensated by the long-run benefit of becoming part of a higher-quality state. However, they do not show explicitly how these improvements in state quality are driven by selection through conflict. They use a dataset with a bigger geographical scope (all of Europe) but lower resolution: they use urban population from \cite{bairoch1988} as a proxy for development, which is available for fewer cities and fewer years.

@@

While this result may seem like evidence for omitted variable bias in the baseline estimate, that explanation is ultimately not plausible. To cause bias in Column (1) conflict incidents would need to be correlated both with the post-$Conquest$ dummy and with construction --- however, we can see that the latter is not the case: the coefficients on $Conflict$ in Columns (3) and (4) are tiny and not significant by any stretch of the imagination.\footnote
{
The coefficients on $Conflict$ in Table \ref{tab:controls_50y} are conditional correlations. Simply computing the correlation coefficient between the dummies for conflict incidents and for construction yields a value of $-0.01$.
} 
Rather, I suggest that the estimates for $\theta_{conquest}$ are different because they are estimated on different samples. More than two thirds of observations are lost between Columns (2) and (3) because of missing values in the $Conflict$ variable. Figure \ref{fig:conflict_map} shows that the remaining observations follow quite a different distribution across space --- if the spatial composition of the smaller sample was the same, the entire map would be a uniform shade. Instead we see big discrepancies: the west, the south, and Brandenburg retain 30\% to 60\% of their observations, while many northern and eastern regions all but vanish from the sample.

If we accept that the difference between the $\theta_{conquest}$ coefficients in (1) and (3) is due to the different spatial distribution of the sample, that has the interesting implication of spatially heterogeneous treatment effects. The null effect in (1) could be an average of the negative effect in the southwest and a hypothetical positive effect in the northeast. Outsiders could play a role, in a similar vein to \cite{levine2021, levine2022}'s theory --- the south and the west are bordered by the Habsburg Empire and France, two stable major powers, while the northeast is somewhat sheltered by the North Sea, the Baltic Sea, and the territory that comprises 1937 Poland, which was fragmented throughout much of its history.\footnote
{
Of course, England and Sweden would attack via the North and Baltic seas, but it seems a reasonable assumption that competition for territory would be more intense between states that share a land border.
} 
\cite{cervellati2022} BLANK. I leave this for future research to explore.
